---
title: "Differential transcript expression and differential transcript usage using Salmon, edgeR v4 and the human T2T reference"
author: "Xueyi Dong"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    code_folding: show
    toc_float: true
    number_sections: true
---

<style type="text/css">
.main-container {
  max-width: 1800px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
options(width = 80, digits = 3, message = FALSE)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
knitr::opts_knit$set(root.dir = "/vast/scratch/users/dong.x/docker/NatProt_LUAD")
```

# Introduction

This document composes different stages of this transcript-level analysis protocol and displays the expected outputs when running on the test data we provided.

Except for the packages required by the protocol, we use an addition package `tictoc` to calculate the time required by each stage that runs R code. The elapsed time of each R-based stage are printed at the end of each stage.

All code shown in this document is imported from scripts for each stage. Code in stage 0 and Stage 1 are UNIX commands and should be run in the terminal, while code in stage 2-6 are R code and should be run in R.

# Stage 0: preparation of transcriptome index

This stage should be run in bash.

```{bash stage0, file="workflow/0_prepare_index.sh", echo=TRUE, eval=FALSE}
```

# Stage 1: transcript quantification (quasi-mapping/selective alignment)

This stage should be run in bash.

```{bash stage1, file="workflow/1_salmon_quantify.sh", echo=TRUE, eval=FALSE}
```

# Stage 2: genes and transcripts annotation

```{r tic2}
library(tictoc)
tic()
```

```{r stage2, file="workflow/2_prepare_annotation.R", echo=TRUE, eval=TRUE}
```

```{r toc2}
toc()
```


# Stage 3: transcript-level count data preprocessing

```{r tic3}
tic()
```

```{r stage3, file="workflow/3_count_preprocess.R", echo=TRUE, eval=TRUE, cache=FALSE}
```

```{r toc3}
toc()
```

# Stage 4: generalized linear model fitting

```{r tic4}
tic()
```

```{r stage4, file="workflow/4_fit_model.R", echo=TRUE, eval=TRUE, cache=FALSE}
```

```{r toc4}
toc()
```

# Stage 5: differential expression analysis

```{r tic5}
tic()
```

```{r stage5, file="workflow/5_differential_expression.R", echo=TRUE, eval=TRUE, cache=FALSE}
```

```{r toc5}
toc()
```

# Stage 6: differential usage analysis

```{r lib, echo=FALSE, eval=TRUE}
library(limma)
library(edgeR)
library(RColorBrewer)
```

```{r tic6}
tic()
```

```{r stage6, file="workflow/6_differential_transcript_usage.R", echo=TRUE, eval=TRUE, cache=FALSE, message = FALSE}

```

```{r toc6}
toc()
```

# Session information

The R session information including version information of used packages is shown below.

```{r}
sessionInfo()
```

